Cluster Dynamics with HETDEX at z < 0.5 - I: Simulated Performance, Mass Distribution and Limits

### 2016-03-06 ms.tex
more comments from casey

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[S. Boada et al.]
{\parbox{\textwidth}{Steven Boada,$^{1}$\thanks{E-mail: boada@physics.tamu.edu}
C.~Papovich,$^{1}$
R.~Wechsler,$^{2,3}$
T. S.~Li,$^{1}$
K.~Gebhardt,$^{5}$
E.~Rozo,$^{2,4}$
E. S.~Rykoff,$^{2}$}\vspace{0.4cm}\
\\
\parbox{\textwidth}{$^{1}$George P.\ and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX, 77843-4242, USA\\
$^{2}$Kavli Institute for Particle Astrophysics and Cosmology, Department of Physics, Stanford University, Stanford, CA 94305, USA\\
$^{3}$Department of Particle Physics and Astrophysics, SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA\\
$^{4}$Department of Physics, University of Arizona, Tucson, AZ 85721, USA\\
$^{5}$Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}}

@@

\begin{abstract}
\noindent
The study of clusters of galaxies has been argued to be a very effective way to measure cosmological parameters, including measuring dark energy and testing models of gravity. The Hobby Eberly Telescope Dark Energy Experiment (HETDEX) will observe many hundreds of square degrees, covering a large sample of galaxy clusters out to $z = 0.5$ based on their optical spectra ($3500-5500\AAA$). The spectra will provide important measures of the clusters dynamics and may enable constraints on cosmological parameters, but only if the measurements provide accurate estimates of the total cluster masses. We have carried out a study to investigate the ability of HETDEX to recover accurate galaxy cluster masses over a wide range of masses and redshifts. We used a detailed mock galaxy catalog and present mock observations of two different scenarios: (1) We targeted individual galaxy clusters to investigate the recovery of parameters with such observations. (2) We created and evaluated a HETDEX-like selection "function'' of galaxies over a similarly sized portion of the sky and use well adopted techniques to recover the dynamical properties, such as velocity dispersion and mass. Using both observing strategies, we produce cluster mass probability density functions $P(X|M,z)$, which can be used to determine the probability that a galaxy cluster of given mass (M), located at redshift ($z$) determined using observable parameter (X). 
\end{abstract}

\section{INTRODUCTION}
Our ability to perform precision cosmology with clusters of galaxies has reached a critical turning point. The widely accepted $\Lambda$CDM model of cosmology makes explicit predictions about the number and masses of galaxy clusters throughout the universe. However, connecting these predictions to a set of, sufficiently large in size, observed clusters remains a principal problem. Specifically, the largest threat to modern, precision, cluster cosmology is not the identification of large numbers of clusters (the total number of clusters known is only going up) but the accurate recovery of galaxy cluster mass \citeeg{Sehgal2011,Plank2014, Bocquet2015}.

As mass is not a direct observable, a lot of work is underway to characterize galaxy cluster masses with an observable feature of galaxy clusters. The goal is to constrain, as best possible, $P(X|M,z)$ the probability ($P$) that a galaxy cluster of given mass ($M$), located at redshift ($z$) determined using observable parameter ($X$). The observable parameter most commonly observed X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014, Mantz2015}, cosmic microwave background observations \citeeg{Vanderlinde2010, Sehgal2011} using the Sunyaev-Zel'dovich effect (SZE; \citealt{Sunyaev1972}) or optical studies \citeeg{Rozo2010, Rozo2015} of richness \citeeg{Abell1958, Rykoff2012} or galaxy velocity dispersions \citeeg{Ruel2014, Sifon2015}.

@@

While a large portion of the $\sim10^6$ interloping [\ion{O}{ii}] galaxies will be field (not associated with a bound structure) galaxies, the large area covered by HETDEX is expected to contain as many as 100 Virgo-sized ($M_{dyn}\sim 10^{15}$ \msol) clusters at $z<0.5$ (\editorial{citation?}). The near-complete spectroscopic coverage allows an unprecedentedly detailed look at a very large number of clusters ranging from group scales to the very massive. In addition to the recovery of accurate dynamical masses, detailed investigations of the of dynamical state of the clusters is possible.

Connecting the dynamical properties derived from spectroscopy to the properties inferred from other studies insures the greatest impact on future work. HETDEX overlaps with the Sloan Digital Sky Survey (SDSS; \citealt{Blanton2001a}), SDSS stripe 82 \citep{Annis2014}, the Dark Energy Survey (DES; \citealt{DES2005}), and the upcoming DECam/IRAC Galaxy Environment Survey (DIRGES; PI: Papovich, C. Papovich \etal\ in preparation). \editorial{SHELA and others? Would be good to have a whole list of different things and different wavelengths.} While the potential dataset is very rich, two large issues remain.

@@

\editorial{The wording here needs to be updated when the time comes.} This work plans to address these concerns in the following ways. We use simulated observations which target individual galaxy clusters to investigate the recovery of parameters with such observations. Secondly, we create and evaluate a HETDEX like selection ``function'' of galaxies over a similarly large portion of the sky and use well adopted techniques to recover the dynamical properties, such as velocity dispersion and mass. Each observation strategy will further be constrained with ``ideal'' and ``realistic'' knowledge. Ideal knowledge assumes that we know which individual galaxy is assigned to which cluster. With realistic knowledge this is unknown and must be determined prior to the estimation of the cluster properties. Both of these strategies will better allow future work to predict the number and types of galaxy clusters which should be observed with VIRUS during both the HETDEX survey portion and through targeted follow up observations.

@@

We investigate the accuracy of the halo mass distribution by comparing the cumulative number density of halos above a mass ($M_{200c}$) threshold to the halo mass function (HMF) of \cite{Tinker2008}. We calculate the HMF is calculated at central redshifts of 0.1, 0.2, and 0.4 using {\tt HMFcalc} \citep{Murray2013} and compared to galaxies in a redshift window of $\Delta z\pm0.01$. We find a very good agreement between the predicted HMF and the observed distribution of clusters.

@@

To assign an [\ion{O}{ii}] luminosity to each galaxy in our catalog we place the catalog galaxies on the same CMD and select all SDSS galaxies in a small 2D ($M_r$, $g-r$) bin around the galaxy. We extract all of the SDSS galaxies inside that bin and create a histogram of their [\ion{O}{ii}] luminosities. Using a slice sampling technique \citep{Neal1997} we assign the catalog galaxy an [\ion{O}{ii}] luminosity based on the distribution of SDSS galaxies extracted. For catalog galaxies which are placed on the CMD near no, or very few ($1\leq n<10$) galaxies we assign it zero [\ion{O}{ii}] luminosity or the mean luminosity, respectively.

The right panel of Figure~\ref{fig: oii sdss} shows the CMD of all SDSS galaxies. Two potential catalog galaxies are also placed on the CMD ($M_r, g-r = -17.7,~0.49$ and $M_r, g-r = -21.4,~1.24$) and indicated by the two colored boxes. The histograms show in the Figure's right panel show the probability density histograms of the Log [\ion{O}{ii}] luminosity for the SDSS galaxies in the 2D bin around the individual catalog galaxies. We sample the distribution and assign each catalog galaxy an [\ion{O}{ii}] luminosity which is then converted into a flux.

@@

The spectral coverage allows for the detection of [\ion{O}{ii}] ($\lambda\lambda 3727-3729~\AAA$ doublet) emitters to $z\sim 0.5$ and Ca H ($\lambda 3968.5~\AAA$) and K ($\lambda 3933.7~\AAA$) absorption features to $z\sim 0.4$. HETDEX is expected to detect sources with continuum brighter than $\sdssg=22$ mag, and emission line strengths above $3.5\times10^{-17}$ \ergscm.

In this work we consider two separate observing strategies and two detection limits. The two observing strategies are targeted and survey-like. The targeted observations use ``direct'' observations where each cluster is targeted individual, and every cluster member galaxy is assumed to be observed. The survey observations mimic the HETDEX observation pattern across the sky, where no cluster is directly targeted and not all cluster member galaxies are observed. In both cases we use two detection thresholds. ``Perfect'' observations assume no detection threshold, if a cluster member galaxy is observed, it is also detected. ``Realistic'' observations assume the HETDEX-like detection thresholds where a galaxy may be observed but not necessarily detected.

These observations provide four levels of quality with targeted-perfect being the highest and survey-realistic being the lowest.

@@

The line-of-sight velocity dispersion (LOSVD) is calculated using a method of maximum likelihood following \cite{Walker2006}. We maximize the probability function 
\begin{equation}
  \label{eq: jointGaussian}
p(\{v_1, ..., v_N\})=\displaystyle\prod_{i=1}^{N}\frac{1}{\sqrt{2\pi(\sigma_i^2+\sigma_p^2)}}\exp\biggl[-\frac{1}{2}\frac{(v_i-\langle u \rangle)^2}{(\sigma_i^2+\sigma_p^2)}\biggr]
\end{equation}
where $\sigma_p$, $\langle\mu\rangle$, and $\sigma_i$ is the LOSVD, the average radial velocity and the error on the individual LOSVs respectively. Using a Monte Carlo Markov Chain (MCMC) sampler ({\tt emcee}\footnote{\url{http://dan.iel.fm/emcee/current/}}; \citealt{Foreman-Mackey2013}) which is based on affine-invariant ensemble sampler (see \citealt{Goodman2010} for details on affine-invariant samplers). We draw twenty thousand samples from the posterior probability distribution. Simple priors, $\langle\mu\rangle$ lies between the maximum and minimum LOSV and $0< \sigma_p < 1400$ \kms, are used. When the full distribution of LOSVDs are not used, the final LOSVD is quoted as the median value of the posterior probability distribution with 68\% error bars defined as the 16th and 84th percentiles of the same distribution.

@@

\subsection{Dynamical Mass}\label{sec: mass}
Recently, the relationship between the LOSVD and dynamical mass has been the focus of several studies \citeeg{Evrard2008, Saro2013, Sifon2013, VanderBurg2014}, and a best fitting relationship for the mass enclosed by $r_{200c}$ of the form
\begin{equation}\label{eq:power law}
	M_{200c} = \frac{10^{15}}{h(z)} \bigg{(}\frac{\sigma_{1D}}{A_{1D}} \bigg{)}^{1/\alpha} \Msol
\end{equation}
with $A_{1D} = 1177 \pm 4.2$ \kms\ (\citealt{Munari2013}; referred to as $\sigma_{15}$ in \citealt{Evrard2008} and other works), $\alpha = 1/3$, $h(z) = H(z)/100$, and $\sigma_{1D}$ is the LOSVD of the velocity tracers (dark matter particles, subhalos or galaxies).

@@

\subsubsection{Probability Based}\label{sec:probability method}
We begin with the training sample. After selecting the desired features $\vec{x} = \{\sigma, z, ...\}$ we make the joint probability between the true cluster mass (M) and $\vec{x}$. Because $\vec{x}$ can be multidimensional, we rely on the corner plot to visualize the relationship between all of the training features. Figure~\ref{fig: probability corner} shows all of the one (marginalized probability) and two (joint probability) dimensional projections of the posterior probability distributions of the features of the training data.

@@

\begin{figure*}
	\includegraphics[width=0.8\textwidth]{figures/recovery.pdf} 
	\caption{Recovery fractions ($N_{obs}/N_{True}$) of cluster member galaxies as a function of redshift and mass for the targeted and survey observing strategies. The solide lines are the median values and the shaded regions represent the 68\% scatter. The significant decline in galaxies observed with the survey strategy is due to gaps in the VIRUS IFU.} \label{fig: recovery} 
\end{figure*}

For the clusters making up the \emph{test} sample the mass is unknown (it is what we are trying to predict) but the other features are known. To determine the mass probability distribution of a test cluster, $P(M)$ we combine the conditional probability distribution, $P(M|\vec{x})$, created previously with the probability distribution of $\sigma$ through Equation~\ref{eq: Pm}.
\begin{equation}\label{eq: Pm}
	P(M) = \int P(M|\vec{x}) P(\sigma) d\sigma
\end{equation}
The expected mass is determined by integrating the mass probability, $P(M)$ over all mass. This becomes our ``predicted'' mass, $\langle M\rangle$.
\begin{equation}\label{eq: expected mass}
	\langle M\rangle = \int M^\prime P(M^\prime)dM^\prime
\end{equation}
The confidence interval associated with this prediction can be estimated two ways. First, by calculating the variance about the expected mass through
\begin{equation}\label{eq: variance}
	V = \int (M^\prime - \langle M\rangle)^2 P(M^\prime)dM^\prime
\end{equation}
or by drawing many samples from $P(M)$ and calculating the values at the 16th and 84th percentile. In practice we find that both methods produce similar results for a large number of trials. Therefore, we quote predicted masses as the most probable mass given by Equation~\ref{eq: expected mass} and associated 68\% error estimated through Equation~\ref{eq: variance}.

@@

\subsection{Recovery of Cluster Members}
As discussed in Section \ref{sec: observations} the observational constraints place limits on the total number fo clusters member galaxies expected to be recovered. Knowing these limits will provide important information for potential future follow up or targeted observations.

Figure \ref{fig: recovery} shows the recovery fraction of member galaxies, the number of observed galaxies divided by the number of actual galaxies ($N_{obs}/N_{True}$), as function of both redshift and cluster mass. It is important to note that if few than five member galaxies are observed the cluster is not considered detected, and is excluded from this figure. As expected, the targeted observing strategy where the individual clusters are targeted through several dithers to ensure near complete coverage, performs significantly better than the survey observing strategy across all redshifts and cluster masses.

For the clusters recovered as a function of redshift, there are two effects at work. The decrease in recovery fraction with increasing redshift is a magnitude effect. If, instead of being limited to $\sdssg= 22$ mag, the observations where limited by absolute magnitude, the sharp downward trend disappears. The second key feature is the strong decline in clusters recovered from survey observations. This is due to gaps in the VIRUS IFU. The median recovery fraction in survey observations is almost exactly 4.5 times less than the targeted median recovery fraction. As the total filling factor of the survey increases the two lines will converge.

The recovery rate as a function of cluster mass, right panel of Figure \ref{fig: recovery}, shows that of the the low mass clusters we detect ($N_{obs} >5$), observe the majority of the galaxies. This also shows a rapid decrease in the detection fraction, which can again be explained by considering absolute magnitudes instead of apparent magnitudes. Just as before, if the survey was limited by absolute magnitude, we find a much more consistent detection fraction.

\subsection{Mass estimates}
In this section we discuss the how accurately we are able to reproduce the true cluster mass from a set of observations. We report on two methods the probability based approach (Section \ref{sec:probability method}) and the ML based method (Section \ref{sec:machine learning method}). For each method we consider both targeted and HETDEX-like observing strategies.

In both figures, we include the cluster masses recovered through the power law scaling relation given in Equation \ref{eq:power law} for both the targeted and survey observations. It should serve as baseline to compare the probability based and ML cluster mass recovery methods. And, while there are many possible metrics to evaluate performance, we compute two: the median absolute error (MAE) and the root mean squared error (RMSE). Both metrics evaluate how closely the ensemble of predicted cluster masses are to the true cluster masses, and in both cases lower numbers are better.

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[S. Boada et al.]
{\parbox{\textwidth}{Steven Boada,$^{1}$\thanks{E-mail: boada@physics.tamu.edu}
C.~Papovich,$^{1}$
R.~Wechsler,$^{2,3}$
T. S.~Li,$^{1}$
K.~Gebhardt,$^{5}$
E.~Rozo,$^{2,4}$
E. S.~Rykoff$^{2}$}\vspace{0.4cm}\
\\
\parbox{\textwidth}{$^{1}$George P.\ and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX, 77843-4242, USA\\
$^{2}$Kavli Institute for Particle Astrophysics and Cosmology, Department of Physics, Stanford University, Stanford, CA 94305, USA\\
$^{3}$Department of Particle Physics and Astrophysics, SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA\\
$^{4}$Department of Physics, University of Arizona, Tucson, AZ 85721, USA\\
$^{5}$Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}}

@@

\begin{abstract}
\noindent
The study of clusters of galaxies has been argued to be a very effective way to measure cosmological parameters, including measuring dark energy and testing models of gravity. The Hobby Eberly Telescope Dark Energy Experiment (HETDEX) will observe many hundreds of square degrees, covering a large sample of galaxy clusters out to $z = 0.5$ based on their optical spectra ($3500-5500\AAA$). The spectra will provide important measures of the clusters dynamics and may enable constraints on cosmological parameters, but only if the measurements provide accurate estimates of the total cluster masses. We have carried out a study to investigate the ability of HETDEX to recover accurate galaxy cluster masses over a wide range of masses and redshifts. We used a detailed mock galaxy catalog and present mock observations of two different scenarios: (1) We targeted individual galaxy clusters to investigate the recovery of parameters with such observations. (2) We created and evaluated a HETDEX-like selection "function'' of galaxies over a similarly sized portion of the sky and use well adopted techniques to recover the dynamical properties, such as velocity dispersion and mass, and benchmark these against observations with perfect knowledge. Using both observing strategies, we produce cluster mass probability density functions $P(X|M,z)$, which can be used to determine the probability that a galaxy cluster of given mass (M), located at redshift ($z$) determined using observable parameter (X). \editorial{This sentence will summarize how well HETDEX will do, and how that might be improved with targeted observations.}
\end{abstract}

\section{INTRODUCTION}
\editorial{The introduction is now much longer because I talked about the cons to the different cluster mass recovery methods. I don't much care for really long introductions, so if there is anything that could by cut to shorten it, I would be ok.}
Our ability to perform precision cosmology with clusters of galaxies has reached a critical turning point. The widely accepted $\Lambda$CDM model of cosmology makes explicit predictions about the number and masses of galaxy clusters throughout the universe. However, connecting these predictions to a set of, sufficiently large in size, observed clusters remains a principal problem. Specifically, the largest threat to modern, precision, cluster cosmology is not the identification of large numbers of clusters (the total number of clusters known is only going up) but the accurate recovery of galaxy cluster mass \citeeg{Sehgal2011,Plank2014, Bocquet2015}.

As mass is not a direct observable, a lot of work is underway to characterize galaxy cluster masses with an observable feature of galaxy clusters. The goal is to constrain, as best possible, $P(M|x)$ the probability ($P$) that a galaxy cluster of given mass ($M$), located at redshift ($z$) determined using observable parameter ($X$). Generally, cluster mass calibrations fall into two distinct camps, simulation based and direct or statistical calibration.

One could use various simulations to attempt to calibrate this observable--mass relation \citeeg{Vanderlinde2010, Sehgal2011}. However, the primary challenge to this method is the incomplete understanding of the baryonic physics which take place in galaxy cluster environments. While there have been (and continue to be) many improvements in the accuracy and power of simulations it is doubtful that in the coming years they will reach the accuracy level required to the point where the observable--mass relation is dominated only by statistics \citep{Weinberg2013}.

The second broad camp is the direct calibration of cluster masses. This recipe has two distinct but not always independent tracks. The ``direct'' method uses the direct observations of a small set of clusters and then uses known mass estimators, X-ray temperatures and luminosities \citeeg{Mantz2010, Rykoff2014}, cosmic microwave background (CMB) observations \citeeg{Vanderlinde2010, Sehgal2011}, optical richness \citeeg{Abell1958, Rykoff2012} or weak lensing (WL) \citeeg{Rozo2010} as examples, which provide a ``true'' mass. This directly calibrates the observable--mass relation which is then applied to a much larger sample. The complications lie in that the ``true'' masses are in fact estimations, and the methods used to recover these masses are subject to their own limitations. X-ray hydrostatic estimations assume hydrostatic equilibrium \citeeg{Mantz2015} which may only be valid for a very small number and range of cluster masses. The Sunyaev--Zel'dovich Effect (SZE; \citealt{Sunyaev1972}), which uses the up--scattering of CMB photons to estimate cluster masses, provides accurate estimations of mass, but the ability to detect low mass galaxy clusters is currently limited by technology \citeeg{Carlstrom2002a}. WL estimates are, in principle, correct in the mean, but they suffer from signal-to-noise requirements, limiting their usefulness in low mass clusters, and potentially suffer from line-of-sight effects as the effect is sensitive to all mass along the line of sight. Virial mass estimators which determine the cluster mass based on the motions of the member galaxies \citeeg{Ruel2014, Sifon2015} are promising in that it is a direct measurement of the depth of clusters potential well, but suffers from systematics due to cluster formation physics which disrupts the velocity field.

The statistical method of determining galaxy cluster mass relies not on direct measurements of individual clusters but the calibration of observables for the entire sample which correlate with cluster mass. One example is the spatial clustering of the galaxy clusters themselves. See \cite{Weinberg2013} for a comprehensive review. In practice, it will be a combination of the three methods touched on that will provide the most reliable determination of cluster masses.

@@

While a large portion of the $\sim10^6$ interloping [\ion{O}{ii}] galaxies will be field (not associated with a bound structure) galaxies, the large area covered by HETDEX is expected to contain as many as 50 Virgo-sized ($M_{dyn}\sim 10^{15}$ \msol) clusters at $z<0.5$ (\editorial{talk to casey about how I calculated this number}). The near-complete spectroscopic coverage allows an unprecedentedly detailed look at a very large number of clusters ranging from group scales to the very massive. In addition to the recovery of accurate dynamical masses, detailed investigations of the of dynamical state of the clusters is possible.

Connecting the dynamical properties derived from spectroscopy to the properties inferred from other studies insures the greatest impact on future work. HETDEX overlaps with the Sloan Digital Sky Survey (SDSS; \citealt{Blanton2001a}), SDSS stripe 82 \citep{Annis2014}, the Dark Energy Survey (DES; \citealt{DES2005}), and the upcoming DECam/IRAC Galaxy Environment Survey (DIRGES; PI: Papovich, C. Papovich \etal\ in preparation), the The \emph{Spitzer}/HETDEX Exploratory Large-Area Survey (SHELA; Papovich \etal\ 2016, ApJS, submitted), and The \emph{Spitzer} IRAC Equatorial Survey (SpIES; Timlin \etal\ 2016, ApJ, submitted). While the potential dataset is very rich, two large issues remain.

@@

\editorial{The wording here needs to be updated when we settle on the verbage for everything else.} This work plans to address these concerns in the following ways. We use simulated observations which target individual galaxy clusters to investigate the recovery of parameters with such observations. Secondly, we create and evaluate a HETDEX like selection ``function'' of galaxies over a similarly large portion of the sky and use well adopted techniques to recover the dynamical properties, such as velocity dispersion and mass. Each observation strategy will further be constrained with ``ideal'' and ``realistic'' knowledge. Ideal knowledge assumes that we know which individual galaxy is assigned to which cluster. With realistic knowledge this is unknown and must be determined prior to the estimation of the cluster properties. Both of these strategies will better allow future work to predict the number and types of galaxy clusters which should be observed with VIRUS during both the HETDEX survey portion and through targeted follow up observations.

@@

We investigate the accuracy of the halo mass distribution by comparing the cumulative number density of halos above a mass ($M_{200c}$) threshold to the halo mass function (HMF) of \cite{Tinker2008}. We calculate the HMF is calculated at central redshifts of 0.1, 0.2, and 0.4 using {\tt HMFcalc} \citep{Murray2013} and compared to galaxies in a redshift window of $\Delta z\pm0.01$. We find a very good agreement between the predicted HMF and the observed distribution of clusters. \editorial{Could show a figure comparing the Buzzard catalog to a plot of the HMF for a couple of redshifts. Might be helpful, might not matter.}

@@

To assign an [\ion{O}{ii}] luminosity to each galaxy in our catalog we place the catalog galaxies on the same CMD and select all SDSS galaxies in a small 2D ($M_r$, $g-r$) bin around the galaxy. We extract all of the SDSS galaxies inside that bin and create a histogram of their [\ion{O}{ii}] luminosities. Using a slice sampling technique \citep{Neal1997} we assign the catalog galaxy an [\ion{O}{ii}] luminosity based on the distribution of SDSS galaxies extracted. For catalog galaxies which are placed on the CMD near no, or very few ($1\leq n<10$) galaxies we assign it zero [\ion{O}{ii}] luminosity or the mean luminosity, respectively. For the galaxies in the Buzzard catalog, 1.3\% of the galaxies brighter than $\sdssg = 22$ mag (HETDEX's detection limit) have exactly zero [\ion{O}{ii}] luminosity. Of the 1.3\%, only 3.3\% (0.05\% of all galaxies) have $g-r < 1.5$ suggesting the galaxies which are assigned zero [\ion{O}{ii}] luminosity are not significantly biasing our simulations.

The right panel of Figure~\ref{fig: oii sdss} shows the CMD of all SDSS galaxies. Two potential catalog galaxies are also placed on the CMD ($M_r, g-r = -17.7,~0.49$ and $M_r, g-r = -21.4,~1.24$) and indicated by the two colored boxes. The histograms show in the Figure's right panel show the probability density histograms of the Log [\ion{O}{ii}] luminosity for the SDSS galaxies in the 2D bin around the individual catalog galaxies. We sample the distribution and assign each catalog galaxy an [\ion{O}{ii}] luminosity which is then converted into a flux. \editorial{We could add a plot like the corner plot to show the multi-dimensionality of this process.}

@@

The spectral coverage allows for the detection of [\ion{O}{ii}] ($\lambda\lambda 3727,3729~\AAA$ doublet) emitters to $z\sim 0.5$ and Ca H ($\lambda 3968.5~\AAA$) and K ($\lambda 3933.7~\AAA$) absorption features to $z\sim 0.4$. HETDEX is expected to detect sources with continuum brighter than $\sdssg=22$ mag, and emission line strengths above $3.5\times10^{-17}$ \ergscm.

In this work we consider two separate observing strategies, targeted and survey-like. The targeted observations use ``direct'' observations where each cluster is targeted individual, and every cluster member galaxy is assumed to be observed. The survey observations mimic the HETDEX observation pattern across the sky, where no cluster is directly targeted and not all cluster member galaxies are observed. Both of these observations have HETDEX-like galaxy detection thresholds, for comparison we also include a set of targeted observations with ``perfect'' knowledge. ``Perfect'' observations assume no detection threshold, if a cluster member galaxy is observed, it is also detected. This provides an important best-case scenario, and differs from the true cluster properties because the recovered cluster properties are still calculated. These observations provide three levels of quality with ``perfect knowledge'' being the highest and survey being the lowest.

@@

The line-of-sight velocity dispersion (LOSVD) is calculated using a method of maximum likelihood following \cite{Walker2006}. We maximize the probability function 
\begin{equation}
  \label{eq: jointGaussian}
p(\{v_1, ..., v_N\})=\displaystyle\prod_{i=1}^{N}\frac{1}{\sqrt{2\pi(\sigma_i^2+\sigma_p^2)}}\exp\biggl[-\frac{1}{2}\frac{(v_i-\langle u \rangle)^2}{(\sigma_i^2+\sigma_p^2)}\biggr]
\end{equation}
where $\sigma_p$, $\langle\mu\rangle$, and $\sigma_i$ is the LOSVD, the average radial velocity and the error on the individual LOSVs respectively. Using a Monte Carlo Markov Chain (MCMC) sampler ({\tt emcee}\footnote{\url{http://dan.iel.fm/emcee/current/}}; \citealt{Foreman-Mackey2013}) which is based on affine-invariant ensemble sampler (see \citealt{Goodman2010} for details on affine-invariant samplers). We draw twenty thousand samples from the posterior probability distribution using simple priors, $\langle\mu\rangle$ lies between the maximum and minimum LOSV and $0< \sigma_p < 1400$ \kms. When the full distribution of LOSVDs are not used, the final LOSVD is quoted as the median value of the posterior probability distribution with 68\% error bars defined as the 16th and 84th percentiles of the same distribution.

@@

\subsection{Dynamical Mass}\label{sec: mass}
The relationship between the LOSVD and dynamical mass has been the focus of several studies \citeeg{Evrard2008, Saro2013, Sifon2013, VanderBurg2014}, and a best fitting relationship for the mass enclosed by $r_{200c}$ of the form
\begin{equation}\label{eq:power law}
	M_{200c} = \frac{10^{15}}{h(z)} \bigg{(}\frac{\sigma_{1D}}{A_{1D}} \bigg{)}^{1/\alpha} \Msol
\end{equation}
with $A_{1D} = 1177 \pm 4.2$ \kms\ (\citealt{Munari2013}; referred to as $\sigma_{15}$ in \citealt{Evrard2008} and other works), $\alpha = 1/3$, $h(z) = H(z)/100$, and $\sigma_{1D}$ is the LOSVD of the velocity tracers (dark matter particles, subhalos or galaxies).

@@

\subsubsection{Probability Based}\label{sec:probability method}
We begin with the training sample. After selecting the desired features $\vec{x} = \{\sigma, z, ...\}$ we make the joint probability between the true cluster mass (M) and $\vec{x}$. Because $\vec{x}$ can be multidimensional, we rely on the ``corner plot'' as an aid to visualize the relationship between all of the training features. Figure~\ref{fig: probability corner} shows all of the one (marginalized probability) and two (joint probability) dimensional projections of the posterior probability distributions of the features of the training data.

@@

\begin{figure*}
	\includegraphics[width=0.8\textwidth]{figures/recovery.pdf} 
	\caption{Recovery fractions ($N_{obs}/N_{True}$) of cluster member galaxies as a function of redshift and true cluster mass for the targeted and survey observing strategies. We have applied HETDEX-like observational limits on the cluster galaxy detection, and require at least five galaxies to be detected for a cluster to be recovered. The solid lines are the median values and the shaded regions represent the 68\% scatter. The significant decline in galaxies observed with the survey strategy is due to gaps in the VIRUS IFU.} \label{fig: recovery} 
\end{figure*}

For the clusters making up the \emph{test} sample the mass is unknown (it is what we are trying to predict) but the other features are known. To determine the mass probability distribution of a test cluster, $P(M)$ we combine the conditional probability distribution, $P(M|\vec{x})$, created previously with the probability distribution of $\sigma$ through Equation~\ref{eq: Pm}.
\begin{equation}\label{eq: Pm}
	P(M) = \int P(M|\vec{x}) P(\sigma) d\sigma
\end{equation}
The expected mass is determined by integrating the mass probability, $P(M)$ over all mass. This becomes our ``predicted'' mass, $\langle M\rangle$.
\begin{equation}\label{eq: expected mass}
	\langle M\rangle = \int M^\prime P(M^\prime)dM^\prime
\end{equation}
The confidence interval associated with this prediction can be estimated two ways. First, by calculating the variance about the expected mass through
\begin{equation}\label{eq: variance}
	\mathrm{Var} = \int (M^\prime - \langle M\rangle)^2 P(M^\prime)dM^\prime
\end{equation}
or by drawing many samples from $P(M)$ and calculating the values at the 16th and 84th percentile. In practice we find that both methods produce similar results for a large number of trials. Therefore, we quote predicted masses as the most probable mass given by Equation~\ref{eq: expected mass} and associated 68\% error estimated through Equation~\ref{eq: variance}.

@@

\subsection{Recovery of Cluster Members}
As discussed in Section \ref{sec: observations} the observational constraints place limits on the total number fo clusters member galaxies expected to be recovered. Knowing these limits will provide important information for potential future follow up or targeted observations. In this section we discuss the number of cluster members recovered for both the targeted and survey observation strategies, taking into account all observational limits.

Figure \ref{fig: recovery} shows the recovery fraction of member galaxies, the number of observed galaxies divided by the number of actual galaxies ($N_{obs}/N_{True}$), as function of both redshift and cluster mass. It is important to note that if fewer than five member galaxies are observed $N_{obs} <5$ the cluster is not considered detected, and is excluded from this figure. As expected, the targeted observing strategy where the individual clusters are targeted through several dithers to ensure near complete coverage, performs significantly better than the survey observing strategy across all redshifts and cluster masses.

For the clusters recovered as a function of redshift, there are two effects at work. The decrease in recovery fraction with increasing redshift is a magnitude effect. We check this by limiting the cluster galaxy detection by absolute magnitude which increases the recovery fraction $>70\%$ at all redshifts, implying the decline is a result of the apparent magnitude cluster galaxy detection threshold. The second key feature is the strong decline in clusters recovered from survey observations. This is due to gaps in the VIRUS IFU. The median recovery fraction in survey observations is almost exactly 4.5 times less than the targeted median recovery fraction. As the total filling factor of the survey increases the two lines will converge.

The recovery rate as a function of cluster mass, right panel of Figure \ref{fig: recovery}, shows that of the the low mass clusters we detect ($N_{obs} >5$), we observe the majority of the galaxies. This also shows a rapid decrease in the detection fraction, which can again be explained by considering absolute magnitudes instead of apparent magnitudes, as above. Also, high mass clusters are rare, so inorder to detect them we must probe a large volume of space. The higher redshift cluster members suffer from the limiting apparent magnitude and suppress the recovery fraction at fixed mass. If we were to limit the survey to $z<0.2$ we find the recovery fraction of clusters, at all masses, increases substantially, and we find a much more consistent detection fraction across masses.

\subsection{Mass estimates}
\begin{table*}
\centering
\caption{Summary of the errors associated with the Targeted and Survey observation strategies, as an ensemble of predictions. See the text for discussion about the MAE and RMSE. Overlap is the percentage of clusters where the true cluster mass is bracketed by the prediction intervals of the predicted mass. See Sections \ref{sec:probability method} and \ref{sec:machine learning method} for a discussion on prediction intervals.}
\begin{tabular}{cccccccc}
	%\hline
	&& \multic{3}{Targeted} & \multic{3}{Survey} \\
	\cline{3-5} \cline{6-8}
	& Input Features & MAE & RMSE & Overlap & MAE & RMSE & Overlap \\
	& & (dex) & (dex) & (\%) & (dex) & (dex) & (\%) \\
	\hline
	\rottext{4}{Prob Based} & Power Law & 0.259 & 0.394 & \nd & 0.243 & 0.357 & \nd \\
	&$\sigma$ & 0.220 & 0.337 & \nd & 0.210 & 0.310 & \nd \\
	&$\sigma, z$ & 0.193 & 0.283 & \nd & 0.184 & 0.271 & \nd \\
	&$\sigma, z, N_{gal}$ & 0.128 & 0.204 & \nd & 0.141 & 0.215 & \nd \\
	\hline
	\rottext{4}{ML Based} & Power Law & 0.258 & 0.394 & \nd & 0.255 & 0.380 & \nd \\
	&$\sigma$ & 0.241 & 0.359 & \nd & 0.223 & 0.342 & \nd \\
	&$\sigma, z$ & 0.193 & 0.282 & \nd & 0.197 & 0.290 & \nd \\
	&$\sigma, z, N_{gal}$ & 0.118 & 0.192 & \nd & 0.140 & 0.221 & \nd \\
	\hline
\end{tabular}
\label{tbl:mass comparisons}
\end{table*}

In this section we discuss the how accurately we are able to reproduce the true cluster mass from a set of observations. We report on two methods the probability based approach (Section \ref{sec:probability method}) and the ML based method (Section \ref{sec:machine learning method}). For each method we consider both targeted and HETDEX-like observing strategies.
